Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-09-04 paper-correlations.tex
Color different clusters on SEM images
almost done

\begin{abstract}
  What have we done?
\end{abstract}

@@

Two ways exist to obtain correlation functions for a given structure at hand:
measure them either experimentally with the help of scattering intensity
\cite{debye1957scattering,li2018direct} or from digital images
\cite{berryman1985measurement,ma2018SS}. None is perfect, as the first approach
suffers from the limitation of CFs that can be obtained this way, while the
second one provides information with limited resolution or/and resolution to
field-of-view ratio \cite{gerke2015universal}. In addition to the resolution
conundrum, the most useful imaging methods such as X-ray computed tomography
(XCT) and scanning electron microscopy (SEM) provide gray-scale images (e.g.,
X-ray attenuation or electron back-scattering distributions) due to their
underlying physical principles and require labeling of constituent phases (or
segmentation) before computation of CFs \cite{samarin2023robust}. If properly
segmented \cite{NNseg}, high-resolution digital images do provide a possibility
to compute any correlation function and, thus, possibility to address numerous
fundamental and practical research problems.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/figa1.jpg}
  \caption[]{A schematic depiction of a binary porous media (pores are shown in
    color) with examples of positive events for surface $F_{ss}$, $F_{sv}$ and
    two-point probability $S_2$ correlation functions. The zoomed in area
    represents the difference between the true ``continuous'' interface in
    between pore and solid phases with pixelized ``digital'' interface emerging
    due to limited resolution of digital images.}
  \label{fig:functions}
\end{figure}

@@

\section{A brief introduction to correlation functions}
\label{sec:math}
In this section we shall provide major properties of correlation functions
supported by our package; the full list of functions is 
(see \cref{fig:functions} for graphical explanation in addition to definitions below):
\begin{itemize}
\item Two-point probability function $S_2$,
\item Lineal-path function $L_2$,
\item Cluster function $C_2$,
\item Surface-surface function $F_{ss}$,
\item Surface-void function $F_{sv}$,
\item Pore size function $P$,
\item Chord length function $p$,
\item Phase cross-correlation function $\rho_{ij}$.
\end{itemize}
Note that cross-correlation $\rho_{ij}$ is actually not a classical CF, but
technically is a variety of $S_2$ for the case when the end of line segment lie
in different phases. This function is useful for multi-phase structures only, as
for binary media two-point probability CFs are interdependent.

@@

\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image is usually faster.  Both
implementations are contained in \code{Directional} and \code{Map} modules of
the package, correspondingly.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/figa_direction.jpg}
  \caption[]{Available directions to compute directional correlation functions in 2D and 3D cases with line scanning approach.}
  \label{fig:directions}
\end{figure}

@@

For surface-surface and surface-void functions we use an edge detection filter
\cite{samarin2023robust} and then calculate either autocorrelation of the edge
or cross-correlation of the edge and the void phase. The edge is extracted by
convolving an input with a short high-pass filter $H$. It has a width $7$ and
all its coefficients with exception of the central coefficient are inversely
proportional to distance from the center:
\begin{equation}
  \begin{aligned}
    H_{ij} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2}} & \quad i \ne 3, j \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right. \\
    H_{ijk} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2 + (k-3)^3}} & \quad i \ne 3, j \ne 3, k
      \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right.
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $C$ is such that all coefficients of $H'$ sum to zero, and $S$ equals to
$30.45849$ in 2D and $172.96232$ in 3D case. The parameters of the filter were adjusted
based on comparison against exact computations of surface functions on smooth-boundary
sets \cite{postcov23}.

@@

Our library does not have an implementation for lineal-path and chord-length maps, although we
know that such implementations exist \cite{turner2016efficient,Havelka}. The reason is a very
slow computation even on GPUs, plus the resulting map is not for all points on
the image, but within a spherical volume (due to rotation of the 3D image during
lineal-path evaluation). Moreover, linear scan approach is orders of magnitude
faster for $L_2$.

@@

Another function built on top of distance transform is the pore size function
$P(r)$. Reckon that pore size function is a probability distribution
function. In practice, it is better represented as a histogram of pore sizes
found in the input image. Our algorithm for calculating $P(r)$ is as follows:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A$}
    \State $D \gets \mathcal{D}(A)$
    \Comment Apply $\mathcal{D}$ (\cref{eq:distance-transform}) to each element of $A$
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to histogram $H$
    \State \textbf{return} $H$
    \Comment Alternatively just return $D'$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

@@

\subsection{Comparison against analytical solutions}
To test out package we generate a realization of random overlapping disks
(\cref{fig:overlapping-disks}) with dimensions $10000 \times 10000$ and
parameters of radii $R = 40\ pixels$ and $\lambda = 10^{-4}\ pixels^{-2}$ and then compare calculated
correlation functions with their theoretical values.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.45\linewidth]{images/disks.png}
  \includegraphics[width=0.45\linewidth]{images/balls.png}
  \caption[]{Images of overlapping disks and balls used in testing of our library.}
  \label{fig:overlapping-disks}
\end{figure}

@@

Correlation functions of XCT images can be seen on \cref{fig:real-xct} and correlation functions of SEM images are on
\cref{fig:real-sem}. Cross-correlation functions for a multiphase XCT image is on \cref{fig:real-multiphase}.

\begin{table}[!pt]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Sample name & Image type & Dimensions (pixels) & Image resolution ($\mu m$) & $C_{0.5}$\\
    \hline
    Ceramic 1   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9823 \\
    Ceramic 2   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9890 \\
    Ceramic 3   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9712 \\
    Sandstone 1 & SEM &  $1280 \times 869$ & 0.10 & 0.9703 \\
    Sandstone 2 & SEM &  $1280 \times 869$ & 0.05 & 0.9897 \\
    Carbonate 1 & SEM &  $1024 \times 691$ & 0.06 & 0.9721 \\
    \hline
    Soil (multiphase sample) & XCT & $304 \times 304 \times 304$ &
    \textcolor{red}{2.25} & \\
    \hline
  \end{tabular}
  \caption{Summary of samples used for calculation of correlation functions.}
  \label{tab:samples}
\end{table}

@@

\begin{figure*}[pt]
  \centering
  \subfigure[Sandstone 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone1.png}}
  \hfill
  \subfigure[Sandstone 2]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone2.png}}
  \hfill
  \subfigure[Carbonate 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/carbonate1.png}}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/s2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/c2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/l2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/surf2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/surfvoid-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/pore-size-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/pore-size-sem.png}
  \caption[]{SEM images and plots of correlation functions for these images.}
  \label{fig:real-sem}
\end{figure*}

@@

@@

\section{Discussion and outlook}
In the light of the end of the previous section it is important to clarify that available
(video) RAM or speed of computations are not the only considerations behind choosing between full correlation map
and directional linear scan. Directional CFs are extremely useful in describing or reconstructing
anisotropic structures \cite{jiao2014chawla,EPL1,tensorPRE}. At the same time, they also much easier to
work with and good candidates for data compression and feature extraction \cite{KarsaninaEJSS}. On the other hand,
full map for $S_2$ allows computing ensemble averaged two-point probability - in the similar fashion it is
measured with scattering experiments. The usefulness of full correlation maps for other CFs is questionable,
as they are either non-measurable experimentally, e.g., \cite{Gommes_notchord}, or non-applicable for stochastic reconstructions.
The latter refers to our ability to reconstruct almost any structure from its $S_2$ map \cite{fullwood2008}. And while we
have fast and efficient techniques such as phase-recovery to reconstruct the structure from its $S_2$ correlation map, the utilization
of other CFs is complicated \cite{cherkasov2021,cherkasovIC}. Simulated-annealing based methods that operate on the full map are, 
to the best of our knowledge, were not developed to date. This is the reason we did not invest into improvement on existing approaches to compute maps for $L_2$.

The trade-off in using full map or directional CFs is closely related to the information content of correlation
functions \cite{Gommes1,Gommes2,skolnick2021}. Any reduction of the full map into ensemble averaged or
directionally computed CFs significantly decreases the information available. On the other hand, full correlation map is
larger than the image itself and, thus, storing such huge amount of data for already voluminous datasets is not practical.
Vectorization of directional correlation functions is, thus, seems like a natural feature reduction approach while
working with correlation maps. As applied to stochastic reconstructions, similar logic applies -
it seems to be beneficial to increase the number of functions rather than their "volume". Full correlation maps
are usually not available unless the structure is already known; but they may be useful in the form of 2D maps
available from images for 3D reconstructions \cite{cherkasov2021}. Wether reconstructions of, say, full 3D maps
based on 2D maps is a better approach than using directional CFs computed from images for stochastic reconstructions
is an open question. Simulated annealing based methods were developed in the first place due to absence of
methods like phase-recovery for correlation functions other than autocorrelation $S_2$. We believe that our fast and robust
computational framework in the form of \code{CorrelationFunctions.jl} package will facilitate research in these areas.

When to use map and when to use linear scan
  memory
  how fast
  optimization during reconstruction
  +directional CFs for anisotropic media
Reconstructions based on maps - from Cherkasov to SA?
Why we need more functions? Information content
3-point in future - do we need more points?
Or may be we need more CFs? Add topology (Zubov)?
GPU computations are limited by RAM - if we are interested in correlation length of width of cube that can be
put into GPU card memory - we can ensemble by computation in subcubes (possible future upodates)
clusterization on GPUs
Numerous directions or solid angle

\section{Summary}
\label{sec:summary}
Text

>>>>

\begin{abstract}
Correlation functions are becoming one of the major tools for quantification of structural information that is usually represented as material's 2D or 3D image.
In this paper we introduce \code{CorrelationFunctions.jl} open-source package developed in Julia and capable of computing all classical
correlation functions based on imaging input data. Images include both binary and multi-phase representations.
Our code is capable of evaluating two-point probability $S_2$, cluster $C_2$, lineal-path $L_2$, surface-surface $F_ss$,
surface-void $F_sv$, pore-size $P$ and chord-length $p$ functions on both CPU and GPU architectures.
There possible, we presented two types of computations: full correlation map (correlations of each point with other pints on the image, that also allows obtaining ensemble averaged CF) and directional correlation functions (currently in major orthogonal and diagonal directions). 
Such an implementation allowed for the first time to assemble a completely free solution to evaluate correlation functions under any operating system with well documented API. Our package includes automatic tests against analytical solutions that are described in this paper. We measured wall times for all CPU and GPU implementations and as a rule of thumb full correlation maps on GPU are faster than directional scans using CPU due to the usage of fast Fourier transform. However, full maps require more RAM and, thus, are limited to available GPU RAM resources. On the other hand, directional CFs are memory efficient and can be evaluated for huge datasets - this way they are the first candidates for
structural data compression of feature extraction. The package itself is available through Julia package ecosystem and on GitHub 
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also contains documentation and additional helpful resources such as tutorials.
We believe that a single powerful computational tool such as \code{CorrelationFunctions.jl} will significantly facilitate the usage of correlation
functions in numerous areas of structural description and research, as well as in machine learning applications. We also present some example as applied to
ceramic, soil composite and oil-bearing rock samples based on their 3D X-ray tomography and 2D scanning electron microscope images. Finally, we conclude our paper with
discussion of possible ways to further improve presented computational framework.
\end{abstract}

@@

Two ways exist to obtain correlation functions for a given structure at hand:
measure them either experimentally with the help of, for example, scattering intensity
\cite{debye1957scattering,li2018direct} or from digital images
\cite{berryman1985measurement,ma2018SS}. None is perfect, as the first approach
suffers from the limitation of CFs that can be obtained this way, while the
second one provides information with limited resolution or/and resolution to
field-of-view ratio \cite{gerke2015universal}. In addition to the resolution
conundrum, the most useful imaging methods such as X-ray computed tomography
(XCT) and scanning electron microscopy (SEM) provide gray-scale images (e.g.,
X-ray attenuation or electron back-scattering distributions) due to their
underlying physical principles and require labeling of constituent phases (or
segmentation) before computation of CFs \cite{samarin2023robust}. If properly
segmented \cite{NNseg}, high-resolution digital images do provide a possibility
to compute any correlation function and, thus, possibility to address numerous
fundamental and practical research problems.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/cf_1.png}
  \caption[]{A schematic representation of a binary porous media (pores are shaded) 
  with examples of CFs computations (all classical correlation functions are presented
  according to the legend on the right side). Computations are results of individual
  events (producing either 0 or 1) ensemble averaged over the image, direction or represented as full correlation map.}
  \label{fig:functions}
\end{figure}

@@

\section{A brief primer on correlation functions}
\label{sec:math}
In this section we shall provide major properties of correlation functions
supported by our package; the full list of functions is 
(see \cref{fig:functions} for graphical explanation in addition to definitions below):
\begin{itemize}
\item Two-point probability function $S_2$,
\item Lineal-path function $L_2$,
\item Cluster function $C_2$,
\item Surface-surface function $F_{ss}$,
\item Surface-void function $F_{sv}$,
\item Pore size function $P$,
\item Chord length function $p$,
\item Phase cross-correlation function $\rho_{ij}$.
\end{itemize}
Note that cross-correlation $\rho_{ij}$ is actually not a classical CF, but
technically is a variety of $S_2$ for the case when the end of line segment lie
in different phases. This function is useful for multi-phase structures only, as
for binary media two-point probability CFs are interdependent.

@@

\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image is usually faster, as it leverages on Fast Fourier Transform computational speed.  Both
implementations are contained in \code{Directional} and \code{Map} modules of
the package, correspondingly.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\linewidth]{images/cf_diag.png}
  \caption[]{Available directions to compute directional correlation functions in 2D and 3D cases with line scanning approach.}
  \label{fig:directions}
\end{figure}

@@

For surface-surface and surface-void functions we use an edge detection filter
\cite{samarin2023robust} and then calculate either autocorrelation of the edge
or cross-correlation of the edge and the void phase. The edge is extracted by
convolving an input with a short high-pass filter $H$. It has a width $7$ and
all its coefficients with exception of the central coefficient are inversely
proportional to distance from the center:
\begin{equation}
  \begin{aligned}
    H_{ij} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2}} & \quad i \ne 3, j \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right. \\
    H_{ijk} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2 + (k-3)^3}} & \quad i \ne 3, j \ne 3, k
      \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right.
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $C$ is such that all coefficients of $H$ sum to zero, and $S$ equals to
$30.45849$ in 2D and $172.96232$ in 3D case. The parameters of the filter were adjusted
based on comparison against exact computations of surface functions on smooth-boundary
sets \cite{postcov23}.

@@

Our library does not have an implementation for lineal-path and chord-length maps, although we
know that such implementations exist \cite{turner2016efficient,Havelka}. The reason is a very
slow computation even on GPUs or the resulting map is not for all points on
the image, but within a spherical volume (due to rotation of the 3D image during
lineal-path evaluation). Moreover, linear scan approach is orders of magnitude
faster for $L_2$.

@@

Another function built on top of distance transform is the pore size function
$P(r)$. Reckon that pore size function is a probability distribution
function. In practice, it is conveniently represented as a histogram of pore sizes
found in the input image. Our algorithm for calculating $P(r)$ is as follows:
\begin{algorithmic}[1]
  \Procedure{$P$}{$A$}
    \State $D \gets \mathcal{D}(A)$
    \Comment Apply $\mathcal{D}$ (\cref{eq:distance-transform}) to each element of $A$
    \State $D' = \{ x \in D: x \ne 0\}$
    \Comment Remove zeros
    \State Bin values in $D'$ to histogram $H$
    \State \textbf{return} $H$
    \Comment Alternatively just return $D'$
  \EndProcedure
\end{algorithmic}
Algorithmic efficiency is $O(n)$ where $n$ is the number of elements in the
input array.

@@

\subsection{Comparison against analytical solutions}
To test out package we generate a realization of random overlapping disks
(\cref{fig:overlapping-disks}) with dimensions $10000 \times 10000$ and
parameters of radii $R = 40\ pixels$ and $\lambda = 10^{-4}\ pixels^{-2}$ and then compare calculated
correlation functions with their theoretical values.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.45\linewidth]{images/disks.png}
  \includegraphics[width=0.45\linewidth]{images/balls.png}
  \caption[]{Images of overlapping disks and balls used for testing of our package.}
  \label{fig:overlapping-disks}
\end{figure}

@@

\begin{table}[!pt]
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \hline
    Sample name & Image type & Dimensions (pixels) & Image resolution ($\mu m$) & $C_{0.5}$\\
    \hline
    Ceramic 1   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9823 \\
    Ceramic 2   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9890 \\
    Ceramic 3   & XCT & $500 \times 500 \times 500$ & 2.25 & 0.9712 \\
    Sandstone 1 & SEM &  $1280 \times 869$ & 0.10 & 0.9703 \\
    Sandstone 2 & SEM &  $1280 \times 869$ & 0.05 & 0.9897 \\
    Carbonate 1 & SEM &  $1024 \times 691$ & 0.06 & 0.9721 \\
    \hline
    Soil (multiphase sample) & XCT & $304 \times 304 \times 304$ &
    15.0 & - \\
    \hline
  \end{tabular}
  \caption{Summary of samples used for calculation of correlation functions.}
  \label{tab:samples}
\end{table}

At first, we compare the three ceramic images which have somewhat different genesis depending on the mixture
ratio of powder and binder used for firing the material \cite{ceramics}. As observed from \cref{fig:real-xct}, the structure of
all three samples is visually very similar. The lineal path function supports this observation. The $S_2$ functions are
also quite similar except for Sample 2 having higher porosity that results in elevated curve (the tail for two-point probability
as expected fluctuates around porosity squared). Notably, $S_2$ and $C_2$ are almost identical - this indicates that
pore phase represents a single large cluster. Recall that for $S_2$, $C_2$ and $L_2$ functions the value at correlation length
of 0 is the porosity value. Surface functions, on the other hand, feel significant difference between samples in
how pore-solid interface is organized. This is further evidenced on pore-size and chord-length graphs, with Ceramic 2
sample with the highest porosity having larger amount of large pores. Both $P$ and $p$ show very similar behavior,
but for different correlation lengths; such behavior is attributed to quite regular pores shapes in all ceramics samples -
this is evident from much larger differences between pore-size and chord-length for SEM images considered next.

@@

The \cref{fig:real-sem} presents three 2D images of sandstones and carbonate with computed correlation functions. In this example,
unlike for quite isotropic ceramics, we evaluated CFs in the directional manner. The $S_2$ graph immediately provides us
with porosity values. It also highlights severe anisotropy in sandstone SEM images. For 2D images numerous patches of pores are
not connected to each other and, thus, $S_2$ and $C_2$ are very different. Cluster functions show how connectivity decays with
length, also we can easily observe if given pore structure percolates in a given direction - none of the images percolates
(we highlighted the clusters to make this statement easily verifiable with the naked eye). Note that $C_2$ and $L_2$ were computed
with correlations lengths equal to image dimensions explore connectivity - for other CFs this leads to noised curves due to smaller number of sampled points (for example, see correlation functions' graphs in \cite{Euras2012}).
Lineal paths are close for sandstones (with lower $L_2$ values for Sandstone 1 where large pores are occluded with clays or grain fragments),
but for carbonate with much smaller apertures of pores it is way below $L_2$ for sandstones. Also, $L_2$ suggests that computations
with correlations lengths beyond where lineal path goes to zero are not very informative. For very irregular pores on 2D SEM images,
$P$ and $p$ functions look much less similar compared to the ceramics case on \cref{fig:real-xct}. For 2D images it is also visually
easy to analyze full CF maps, as presented on \cref{fig:real-sem_maps} for Sandstone 1 sample. Full correlation maps show expanded version of directional functions -
asymmetry is immediately evidenced from $S_2$, $F_ss$ and $F_sv$ maps. Moreover, the map for $C_2$ shows higher connectivity in y-direction
as compared to x-axis. The interpretation of full correlation maps is less straightforward, but anisotropy patterns in directions different from
those presented in \cref{fig:directions} can be spotted.

\begin{figure*}[pt]
  \centering
  \subfigure[Sandstone 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone1.png}}
  \hfill
  \subfigure[Sandstone 2]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone2.png}}
  \hfill
  \subfigure[Carbonate 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/carbonate1.png}}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/s2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/c2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/l2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/surf2-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/surfvoid-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.3\linewidth]{samples/pore-size-sem.png}
  \includegraphics[width=0.3\linewidth]{samples/chord-length-sem.png}
  \caption[]{SEM images and plots of correlation functions for these images.}
  \label{fig:real-sem}
\end{figure*}

Finally, we conclude this section with the example of multi-phase soil composite for which we computed cross-correlation functions \cref{fig:real-multiphase}.
The graph on the right side of the figure shows cross-correlations between different pairs of phases (normalized by each phase's ratio). One can easily conclude
that organo-mineral complex (phase 2) is mainly located on pore (phase 1) walls. Heavy minerals (phase 4) do not cross-correlate with pores, as
they are mainly lie within the mineral solid phase (phase 3) - they do show high cross-correlations. Other CFs can be easily computed for each separate phase,
while $F_sv$ can potentially be extended beyond void (or pores) and be represented by any other phase or their superposition.

@@

This section clearly demonstrated, that \code{CorrelationFunctions.jl} can be utilized in analysis of any 2D and 3D images, including structures
consisting of numerous phases.

@@

\section{Discussion and outlook}
In the light of the end of the previous section it is important to clarify that available
(video) RAM or speed of computations are not the only considerations behind choosing between full correlation map
and directional linear scans. Directional CFs are extremely useful in describing or reconstructing
anisotropic structures \cite{jiao2014chawla,EPL1,tensorPRE}. At the same time, they also much easier to
work with and good candidates for data compression and feature extraction \cite{KarsaninaEJSS}. On the other hand,
full map for $S_2$ allows computing ensemble averaged two-point probability - in the similar fashion it is
measured with scattering experiments. The usefulness of full correlation maps for other CFs is questionable,
as they are either non-measurable experimentally, e.g., \cite{Gommes_notchord}, or non-applicable for stochastic reconstructions.
The latter refers to our ability to reconstruct almost any structure from its $S_2$ map \cite{fullwood2008}. And while we
have fast and efficient techniques such as phase-recovery to reconstruct the structure from its $S_2$ correlation map, the utilization
of other CFs is complicated \cite{cherkasov2021,cherkasovIC}. Simulated-annealing based methods that operate on the full map are, 
to the best of our knowledge, were not developed to date. This is the reason we did not invest into improvement on existing approaches to compute maps for $L_2$.

The trade-off in using full map or directional CFs is closely related to the information content of correlation
functions \cite{Gommes1,Gommes2,skolnick2021}. Any reduction of the full map into ensemble averaged or
directionally computed CFs significantly decreases the information available. On the other hand, full correlation map is
larger than the image itself and, thus, storing such huge amount of data for already voluminous datasets is not practical.
Vectorization of directional correlation functions is, thus, seems like a natural feature reduction approach while
working with correlation maps. As applied to stochastic reconstructions, similar logic applies -
it seems to be beneficial to increase the number of functions rather than their "volume". Full correlation maps
are usually not available unless the structure is already known; but they may be useful in the form of 2D maps
available from images for 3D reconstructions \cite{cherkasov2021}. Wether reconstructions of, say, full 3D maps
based on 2D maps is a better approach than using directional CFs computed from images for stochastic reconstructions
is an open question. Simulated annealing based methods were developed in the first place due to absence of
methods like phase-recovery for correlation functions other than autocorrelation $S_2$. We believe that our fast and robust
computational framework in the form of \code{CorrelationFunctions.jl} package will facilitate research in these directions.

While the toolbox of presented package and its API are solid, based on additional research numerous areas for future improvements of \code{CorrelationFunctions.jl}
are possible:
\begin{enumerate}
  \item include novel non-classical correlation functions, e.g., describing topology of phases;
  \item implementing sub-cube based computations of CFs with a given correlation functions with later averaging for larger volume - 
  with such domain decomposition one could compute CFs for larger samples with limited GPU resources;
  \item implementing GPU-based clustering (cluster labeling) and distance-map transform to further speed up computations for $C_2$, chord-length and pore-size functions;
  \item implementing higher order functions, such as 3-point probabilities or even 3-point cluster and other more involved CFs;
  \item increase number of directions for scan approach or implementing calculations within a given solid angle.
\end{enumerate}
We perform steps forward in some directions, whilst others still require additional research to be made to facilitate their implementation
within the package.

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package developed in Julia and capable of computing all classical
correlation functions based on 2D and 3D input images. Our code is capable of evaluating two-point probability, cluster, lineal-path, surface-surface,
surface-void, pore-size and chord-length functions on both CPU and GPU architectures. There possible, we presented two types of computations: full correlation map
(correlations of each point with other pints on the image, that also allows obtaining ensemble averaged CF) and directional correlation functions
(currently in major orthogonal and diagonal directions). Such an implementation allowed for the first time to assemble a completely free solution to
evaluate correlation functions under any operating system with well documented API. Our package includes automatic tests against analytical solutions
that are described in this paper. We measured wall times for all CPU and GPU implementations and as a rule of thumb full correlation maps on GPU are faster
than directional scans using CPU due to the usage of fast Fourier transform. However, full maps require more RAM and, thus, are limited to available GPU RAM
resources. On the other hand, directional CFs are memory efficient and can be evaluated for huge datasets - this way they are the first candidates for
structural data compression of feature extraction. The package itself is available through Julia package ecosystem and on GitHub 
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also contains documentation and additional helpful resources such as tutorials.
We believe that a single powerful computational tool such as \code{CorrelationFunctions.jl} will significantly facilitate the usage of correlation
functions in numerous areas of structural description and research, as well as in machine learning applications.

\section{Acknowledgments}
This research was supported by the Russian Science Foundation grant
19-72-10082.

Collaborative effort of the authors within the FaT iMP (Flow and Transport in
Media with Pores) research group (www.porenetwork.com) and used some of its
software.
